---
title: "week6_exercises_methods_2"
author: "Sigurd Sørensen"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Matrix calculation by hand. 
exercises in the folder for today. 

 - Exercise 4.3 in the book
 - Exercise 1 , 3 and 6 shown as pictures. 
 
```{r}
a1<- matrix(seq(1,4, by =1), ncol = 2)
det(a1)

b1<-matrix(c(0,3,-3,0), ncol=2, byrow = TRUE)
det(b1)

c1<-matrix(c(1,0,3,6), ncol=2, byrow = TRUE)
det(c1)

l1<-matrix(c(3,9,3,-10,3,0,3,0,3,7,3,-1,3,11,3,-1), ncol = 4, byrow = TRUE)
det(l1)
```

```{r}
pacman::p_load(pracma)
a3<-pracma::cross(c(1,2,3),c(1,0,3))

b3<-pracma::cross(c(5,3,7),c(1,0,0))
```
 
 
 
```{r}
a6<-a1
test<-inv(a6)
test
inv(test)

b6<-b1
inv(b6)

f6<-matrix(c(1,0,0,3,6,0,3,5,4), ncol=3, byrow=TRUE)
inv(f6)
```
 
 - 

# Intro to matrix operations in R.

```{r}
pacman::p_load(tidyverse)
```

*Multiplication*
```{r}
matrix1 <- matrix(seq(1,10, by = 1), ncol = 2)

matrix2 <- matrix(seq(1,10, by = 1), nrow = 2)

# 5x2 premultiplied by 2x5 = 5x5
matrix1 %*% matrix2

# 2x5 premultiplied by 5x2 = 2x2 
matrix2 %*% matrix1
```
Another test
```{r}
test1<-matrix(c(0,0,1,0,1,0,1,0,1), nrow=3, byrow=TRUE)

test1 %*% test1 
```


*Transposing*
```{r}
matrix1

# We can transpose using t()
t(matrix1)
```
The first column is now the first row and so forth. 


*Inverse*
```{r}
# solve() is used to find the inverse of a matrix
inv(matrix1)
```
This gives us an error message that saying that we don't have a square matrix and therefore cannot find the inverse. 

A matrix multiplied by its own transposed will always give us a square matrix. 
```{r}
t(matrix1) %*% matrix1
```

```{r}
matrix1 %*% t(matrix1)
```
But as we can see the order still maters. If we take t(X) pre-multiplied by X we will get a square matrix with the size of the orignal X's number of columns. If the order of multiplcation is inversed we will of course get the new matrix with dimensions equal 
to the orignal X's number of rows.

You can test this yourself if you still don't believe me. ;) 


We find the inverse of $(t(X)X)^{-1}$
```{r}
solve(t(matrix1) %*% matrix1)
inv(t(matrix1) %*% matrix1)
```


Another issue that can arise with inversions of matrices. 
```{r}
matrix3 <- matrix( c(1,2,3,6) ,ncol = 2, nrow = 2)
matrix3
```
We see that this matrix is a square matrix so we should be able to find the inverse,
or should we?? Let's try it. 

```{r}
solve(matrix3)
```
It returns the message that is singular. What does that mean? How does it relate to the 
our knowledge of the determinant? 

*Determinant*
We can find the determinant of matrix using the following det() function. 
```{r}
det(matrix3)
```


# Your turn in R. 

## Exercise 1


    1) Simulate random X variables (choose as many as you want) and make a design matrix
    
```{r}
pacman::p_load(tidyverse)
df<-read_csv("winequality.csv")

design <- subset(df, select=c("fixed acidity", "alcohol", "quality"))
design<- design %>% 
    rename(fixaci = 'fixed acidity')
glm(quality ~fixaci, data = design)

v1<-rep(1,length(design$alcohol))
design<-cbind(v1,design)

```
    
    
    2) Decide on your beta values and save it in a beta vector.  
```{r}
beta<-matrix(c(6.192, -0.05187), ncol = 1)
```


    3) Using matrix multiplication find y based on our beta vector and design matrix. 
        3.1) Plot the relationship between y ~ x
        3.2) Does something look odd? 

```{r}
multi<- subset(design, select = c(v1,fixaci))
multi<-data.matrix(multi)
newmatrix<-multi %*% beta

fixedacid<-cbind(newmatrix,design)
fixedacid<- fixedacid %>% 
    rename(quality_pred = newmatrix)

ggplot(data = fixedacid, aes(x=fixaci, y=quality_pred))+
    geom_line()

ggplot(data = fixedacid, aes(x=fixaci, y=quality))+
    geom_line()



```

        
    4) Simulate errors and add it your y vector so that it becomes yhat. 
        4.1) Plot the new y ~ x and the abline corresponding to our beta values. 
        4.2) Is this more realistic than what we observed 3.2? 
        
```{r}
y_hat<-fixedacid$quality-fixedacid$quality_pred
fixedacid<-cbind(fixedacid,y_hat)

ggplot(data = fixedacid, aes(x=fixaci, y=quality_pred+y_hat))+
    geom_line()+
    geom_abline(intercept = beta[1,], slope = beta[2,])

```

    
    5) Using your knowledge of matrix inversion and the general linear model find the corresponding beta values to the yhat ~ x.
        5.1) Did beta coefficinets match the ones you specified in the earlier example and               why/why not? 
        
```{r}
glm(y_hat+quality_pred ~ fixaci, data = fixedacid)
```

        5.2) run normal lm() on yhat ~ x. Does the beta coefficients match? 
        
```{r}
lm(y_hat+quality_pred ~ fixaci, data = fixedacid)
```


        5.3) When is it problematic to find the inverse of a matrix? How does it relate to the error message we sometimes get when running regression models: (isSingular)? 
        
When it doesn't fit in size. 
I'm not sure I quite understand this exercise, but it was fun anyhow...
        
        

## Exercise 2
```{r}
data(mtcars)
mtcars
```

    1) Using matrix multiplication find the beta values of the model: *mpg ~ hp + wt + am*

```{r}

car<-subset(mtcars, select=c(mpg, hp,wt,am))

car<-car %>% 
    mutate(x=hp+wt+am)
int<-rep(1,length(car$x))
car<-cbind(int,car)

mpg<-subset(car, select = mpg)
mpg<-data.matrix(mpg)
car<-data.matrix(car)
car<-subset(car, select = c(int, x))

Beta<-t(car)%*%mpg
Beta
```
*hint:* Design Matrix, and y vector. 

    2) Check if the output is the same as when you use the lm() function

```{r}
cars<-subset(mtcars, select = c(mpg, hp,wt,am))
cars<-cars %>% 
    mutate(x=hp+wt+am)
lm(mpg ~ hp+wt+am,data =cars)
```


```{r}

car2<-subset(mtcars, select=c(mpg, hp,wt,am))

int<-rep(1,length(car2$hp))
car2<-cbind(int,car2)

mpg<-subset(car2, select = mpg)
mpg<-data.matrix(mpg)
car2<-data.matrix(car2)

Beta<-t(car2)%*%mpg
Beta
```


