---
title: "220202-Week1Exercises"
author: "Thomas Steinthal"
date: "2/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, car, MASS, lmtest, tseries, ggfortify)
```

1. Do exercise 1.2 in the book
Sketching a regression model and data: Figure 1.1b shows data corresponding to the fitted line
y = 46.3 + 3.0x with residual standard deviation 3.9, and values of x ranging roughly from 0 to
4%.
(a) Sketch hypothetical data with the same range of x but corresponding to the line y = 30 + 10x
with residual standard deviation 3.9.


```{r}
set.seed(1234)

x1 <- rnorm(17, 2, 2/3)
error <- rnorm(17, 0, 3.9)

y1 <- 30 + 10*x1 + error

m1=lm(y1~x1)
summary(m1)

plot(x1,y1)+abline(30, 10)
```


(b) Sketch hypothetical data with the same range of x but corresponding to the line y = 30 + 10x
with residual standard deviation 10.

```{r}
set.seed(1234)

x2 <- rnorm(17, 2, 2/3)
error_b <- rnorm(17, 0, 10)

y2 <- 30 + 10*x2 + error_b

m2=lm(y2~x2)
summary(m2)

plot(x2,y2)+abline(30, 10)
```


```{r}
df <- read_csv("oscars.csv")

summary(df)
hist(df$Nom)
```

```{r}
wine <- read_csv("winequality.csv")
wine <- wine %>% 
  rename(
    fixacid = `fixed acidity`,
    volacid = `volatile acidity`,
    citacid = `citric acid`,
    ressug = `residual sugar`,
    fresuldio = `free sulfur dioxide`
  )

wine <- wine %>% 
  na.omit(wine)

#Overall data-inspection
ggplot(wine,aes(x=fixacid,y=ressug,colour=quality))+geom_point()+geom_smooth(method=lm,se=FALSE)+labs(title="Data Inspection")

# Basic statistical info of the angle variable, which looks skewed
summary(wine)



#Final boxplot to 
boxplot(fixacid ~ ressug,
          col=c("white","lightgray"),wine)



```

```{r}
m1<- lm(quality ~ fixacid,data=wine)
m2<- lm(quality ~ ressug, data=wine)
m3<- lm(quality ~ alcohol, data=wine)
m4<- lm(quality ~ alcohol*pH, data=wine)

AIC(m1,m2,m3,m4)
```
m4 is found to be the best model



```{r}
#Making a random split of the data
pacman::p_load(caret)
set.seed(340)
trainIndex <- createDataPartition(wine$quality,
                                  p = .8, #we create a partition with 60%
                                  list = FALSE, #not a list but a data.frame
                                  times = 1) #Times we split = 1 time. 
Train <- wine[trainIndex,] #We subset iris_subset by our indexing variable trainIndex
Test <- wine[-trainIndex,] #Everything that is not in our trainIndex/Train will be in Test.
```

```{r}
#Train the model on our Train data set.
model_train <- lm(quality ~ alcohol*pH, data = Train)
```

```{r}
#Predict based on your trained model and Test df partition. type = "reponse" because we're working with logistic regression. This changes the outcome to be probabilities and not log-odds. 
Test$y_hat_prop <-  predict(model_train, Test, type = "response")

Test %>% 
  summarise(y_hat_prop, quality)
```


All in all (even though there maybe should have been put more theoretical work into it, I have now found a model that is able to predict the quality of the wine based on the interactive predictors, alcohol and pH (as I assume them to interact)).



###Exercise 2.3

```{r}
pacman::p_load(stringi)

nam <- read_csv('allnames_clean.csv')

nam<- nam %>% 
  subset(sex == "F")
nam<- nam %>%  mutate(
  laslet = stri_sub(name,-1)
)


namsor<- nam %>% 
  group_by(laslet) %>% 
  summarise_at(vars(X1880:X2010), sum, na.rm = TRUE)


longnam<- namsor %>% 
  pivot_longer(
    cols = X1880:X2010
  )

```



```{r}
pacman::p_load(data.table)

# flip datafram
  t_namsor <- transpose(namsor)

# get row and colnames in order
  colnames(t_namsor) <- rownames(namsor)
  rownames(t_namsor) <- colnames(namsor)
  
# Remove first row
  t_namsor = t_namsor[-c(1),]
  
#plot
  matplot(t_namsor,type=c("l"),pch=1,col=1:26,lty= 1:26,xlab="Years from 1880", ylab="Number of names given")
  #for legends plot together with this
  legend("topleft", legend = 1:26, col=1:26,pch=1,title='Last letter of name')
  
 #By googling I find this way of doing stuff 
nn <- ncol(t_namsor)
layout(matrix(c(1,2),nrow=1), width=c(4,1)) 
par(mar=c(5,4,4,0)) #No margin on the right side
  matplot(t_namsor,type=c("l"),pch=1,col=1:26,lty= 1:26,xlab="Years from 1880", ylab="Number of names given")
par(mar=c(5,0,4,2)) #No margin on the left side
plot(c(0,1),type="n", axes=F, xlab="", ylab="")
legend("center", colnames(t_namsor),col=seq_len(nn),cex=0.5,fill=seq_len(nn),title='Last letter')
```

