---
title: "220209-ExercisesWeek2"
author: "Thomas Steinthal"
date: "9/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse)
```


## Exercise 3.1
 Weighted averages: A survey is conducted in a certain city regarding support for increased
property taxes to fund schools. In this survey, higher taxes are supported by 50% of respondents aged 18–29, 60% of respondents aged 30–44, 40% of respondents aged 45–64, and 30% of respondents aged 65 and up. Assume there is no nonresponse

Suppose the sample includes 200 respondents aged 18–29, 250 aged 30–44, 300 aged 45–64,
and 250 aged 65+. Use the weighted average formula to compute the proportion of respondents
in the sample who support higher taxes.

```{r}
#Groups
n1<-200
n2<-250
n3<-300
n4<-250

p1<-0.5
p2<-0.6
p3<-0.4
p4<-0.3

n<-sum(n1,n2,n3,n4)

sup<- (n1*p1+n2*p2+n3*p3+n4*p4)/n

sup

```
This means that the proportion of the people to support higher taxes are 44.5%

##Exercise 3.3
Probability distributions: Using R, graph probability densities for the normal distribution,
plotting several different curves corresponding to different choices of mean and standard
deviation parameters.
```{r}

ex1<-rbinom(10,size=10, prob= 1/6)
ex2<-rbinom(100,size=10,prob=1/6)
ex3<-rbinom(100,size=20,prob=1/6)
ex4<-rbinom(10000,size=50,prob=1/10)
ex5<-dbinom(1:10, size = 10, prob= 1/6)

plot(density(ex1))
plot(density(ex2))
plot(density(ex3))
plot(density(ex4))

plot(density(ex5))


```


## Exercise 3.4
Using a bar plot in R, graph the Poisson distribution with parameter
3.5.

```{r}
x1<-rpois(100,lambda=3.5)
ex6<-data.frame(x=x1)
ggplot(ex6,aes(x=x))+geom_bar()
```


##Exercise 3.5
Using a bar plot in R, graph the binomial distribution with n = 20 and
p = 0.3.
```{r}

x1<-rbinom(20,size=1,prob=0.3)
ex7<-data.frame(x=x1)
ggplot(ex7,aes(x=x))+geom_bar()

```

## Exercise 3.6
A test is graded from 0 to 50, with an average score of 35 and a standard
deviation of 10. For comparison to other tests, it would be convenient to rescale to a mean of 100 and standard deviation of 15.
(a) Labeling the original test scores as x and the desired rescaled test score as y, come up with a linear transformation, that is, values of a and b so that the rescaled scores yu = a + bx have a mean of 100 and a standard deviation of 15.
(b) What is the range of possible values of this rescaled score 푦?
(c) Plot the line showing 푦 vs. 푥
```{r}
## A
#Mean
x_mean<-35
y_desired_mean<-100

a<-y_desired_mean-1.5*x_mean

x_sd<-10
y_desired_sd<-15

b<-y_desired_sd/x_sd

test<- rnorm(1000, mean=x_mean,sd=x_sd)
trans<-a+test*b

mean(trans)
sd(trans)

hist(test)
hist(trans)
```
```{r}
##b
range(trans)
```

```{r}
##c
df<- data.frame(test=test,trans=trans)

ggplot(df,aes(x=test,y=trans))+geom_line()
```

##Exercise 4.1
A randomized experiment is performed within a survey. 1000 people are contacted. Half the people contacted are promised a $5 incentive to participate, and half are not promised an incentive. The result is a 50% response rate among the treated group and 40% response rate among the control group. Give an estimate and standard error of the average treatment effect

```{r}
tg<-rbinom(1, size = 500, prob = 0.5)
cg<-rbinom(1, size = 500, prob = 0.4)

m<-(500*tg+500*cg)/1000

se_tg<-sqrt(0.5*0.5/500)
se_cg<-sqrt(0.4*0.6/500)

se=sqrt(se_tg^2+se_cg^2)

m
se
```



4.1 4.2 4.3 4.4 & 4.5 
Use the skills you learned: 
Exercise 3.10 & 4.11
Hard: (challenge)
3.2 3.8 & 4.7
