---
title: "220209-ExercisesWeek2"
author: "Thomas Steinthal"
date: "9/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse)
```


## Exercise 3.1
 Weighted averages: A survey is conducted in a certain city regarding support for increased
property taxes to fund schools. In this survey, higher taxes are supported by 50% of respondents aged 18–29, 60% of respondents aged 30–44, 40% of respondents aged 45–64, and 30% of respondents aged 65 and up. Assume there is no nonresponse

Suppose the sample includes 200 respondents aged 18–29, 250 aged 30–44, 300 aged 45–64,
and 250 aged 65+. Use the weighted average formula to compute the proportion of respondents
in the sample who support higher taxes.

```{r}
#Groups
n1<-200
n2<-250
n3<-300
n4<-250

p1<-0.5
p2<-0.6
p3<-0.4
p4<-0.3

n<-sum(n1,n2,n3,n4)

sup<- (n1*p1+n2*p2+n3*p3+n4*p4)/n

sup

```
This means that the proportion of the people to support higher taxes are 44.5%

##Exercise 3.3
Probability distributions: Using R, graph probability densities for the normal distribution,
plotting several different curves corresponding to different choices of mean and standard
deviation parameters.
```{r}

ex1<-rbinom(10,size=10, prob= 1/6)
ex2<-rbinom(100,size=10,prob=1/6)
ex3<-rbinom(100,size=20,prob=1/6)
ex4<-rbinom(10000,size=50,prob=1/10)
ex5<-dbinom(1:10, size = 10, prob= 1/6)

plot(density(ex1))
plot(density(ex2))
plot(density(ex3))
plot(density(ex4))

plot(density(ex5))


```


## Exercise 3.4
Using a bar plot in R, graph the Poisson distribution with parameter
3.5.

```{r}
x1<-rpois(100,lambda=3.5)
ex6<-data.frame(x=x1)
ggplot(ex6,aes(x=x))+geom_bar()
```


##Exercise 3.5
Using a bar plot in R, graph the binomial distribution with n = 20 and
p = 0.3.
```{r}

x1<-rbinom(20,size=1,prob=0.3)
ex7<-data.frame(x=x1)
ggplot(ex7,aes(x=x))+geom_bar()

```

## Exercise 3.6
A test is graded from 0 to 50, with an average score of 35 and a standard
deviation of 10. For comparison to other tests, it would be convenient to rescale to a mean of 100 and standard deviation of 15.
(a) Labeling the original test scores as x and the desired rescaled test score as y, come up with a linear transformation, that is, values of a and b so that the rescaled scores yu = a + bx have a mean of 100 and a standard deviation of 15.
(b) What is the range of possible values of this rescaled score 푦?
(c) Plot the line showing 푦 vs. 푥
```{r}
## A
#Mean
x_mean<-35
y_desired_mean<-100

a<-y_desired_mean-1.5*x_mean

x_sd<-10
y_desired_sd<-15

b<-y_desired_sd/x_sd

test<- rnorm(1000, mean=x_mean,sd=x_sd)
trans<-a+test*b

mean(trans)
sd(trans)

hist(test)
hist(trans)
```
```{r}
##b
range(trans)
```

```{r}
##c
df<- data.frame(test=test,trans=trans)

ggplot(df,aes(x=test,y=trans))+geom_line()
```

##Exercise 4.1
A randomized experiment is performed within a survey. 1000 people are contacted. Half the people contacted are promised a $5 incentive to participate, and half are not promised an incentive. The result is a 50% response rate among the treated group and 40% response rate among the control group. Give an estimate and standard error of the average treatment effect

```{r}
tg<-rbinom(1, size = 500, prob = 0.5)
cg<-rbinom(1, size = 500, prob = 0.4)

m<-(500*tg+500*cg)/1000

se_tg<-sqrt(0.5*0.5/500)
se_cg<-sqrt(0.4*0.6/500)

se=sqrt(se_tg^2+se_cg^2)

m
se
```

##Exercise 4.2 - PROBABLY WRONG
You are designing a survey to estimate the gender gap: the difference in
support for a candidate among men and women. Assuming the respondents are a simple random
sample of the voting population, how many people do you need to poll so that the standard error is less than 5 percentage points?

We assume that there 
```{r}
si<-200
se_m<- sqrt(0.5*0.5/si)
se_f<- sqrt(0.5*0.5/si)

s_diff<-sqrt(se_m^2+se_f^2)
s_diff

```

## Exercise 4.3
 You want to gather data to determine which of two students is a
better basketball shooter. One of them shoots with 30% accuracy and the other is a 40% shooter.
Each student takes 20 shots and you then compare their shooting percentages. What is the
probability that the better shooter makes more shots in this small experiment?

```{r}
gs<-rbinom(1, size = 20, prob = 0.4)
bs<-rbinom(1, size = 20, prob = 0.3)

se_gs<-sqrt(0.4*0.6/20)
se_bs<-sqrt(0.3*0.7/20)

di<-sqrt(se_gs^2+se_bs^2)
di

dbinom()

```

##Exercise 4.4
You want to gather data to determine which of two students is a
better basketball shooter. You plan to have each student take n shots and then compare their
shooting percentages. Roughly how large does n have to be for you to have a good chance of
distinguishing a 30% shooter from a 40% shooter?
```{r}

```


## Exercise 4.5
 Sampling distribution: Download a data file on a topic of interest to you. Read the file into R
and order the data by one of the variables.

a) Use the sample function in R to draw a simple random sample of size 20 from this population.
What is the average value of the variable of interest in your sample?
(b) Repeat this exercise several times to get a sense of the sampling distribution of the sample
mean for this example.

```{r}
wine<- read_csv("winequality.csv")

samp<-sample(wine$alcohol, size = 20, replace = FALSE)
mean(samp)

samp_f <- function(){
  s<- sample(wine$quality, size = 20, replace = FALSE)
  mean(s)
  return(s)
}

samp_f()

samp_dist_of_samp_means <- c()
for (i in 1:1000){
  samp_dist_of_samp_means[i] <- mean(samp_f())
}
hist(samp_dist_of_samp_means)

```


Use the skills you learned: 
Exercise 3.10 & 4.11
Hard: (challenge)
3.2 3.8 & 4.7
